Mitigating Interest Flooding DDoS Attacks in NDN

### 2012-12-19 04b-design-intelligent.tex
More edits to chapters 4a, 4b
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@891 df1a90cd-247b-4364-bdaf-287a704eb499

In order to distinguish between legitimate and malicious Interests, we leverage another unique feature of the stateful forwarding in NDN---namely, the guaranteed symmetric flow of Interest and Data packets. Since a Data packet takes the reverse path of the corresponding Interest packet, a router is guarateed to see if an Interest it forwarded resulted in a matching Data packet. The only exception being if Interest/Data packets were lost along the way due to congestion.
Therefore, intermediate routers can classify all Interests that result in matching Data as legitimate, while the ones that timed out can be marked as malicious.\footnote{Recall that in order to maximize effect of the Interest flooding attack, an adversary expresses a large volume of junk Interests (see Section~\ref{sec:interest flooding}).}  
%Implications of other types of attacks are discussed in Section~\ref{sec:discussion}.}

This timeout-based differentiation method is reactive by nature: one cannot know in advance that an Interest would timeout or bring Data back. However, routers can proactively keep an up-to-date statistics of Interests satisfaction ratios (number of forwarded versus number of satisfied Interests), and use this statistic to determine whether an incoming Interest must be forwarded or dropped. For example, maintaining independent Interest satisfaction ratio statistics for each incoming interface is sufficient to estimate whether an Interest received from a neighboring interface will result in matching Data or a timeout. Statistics can also be kept at finer granularities such as per outgoing interface, per name prefix, etc. that can further improve the estimates. A router's goal is to penalize malicious Interests as soon as possible -- negative statistics should build up fast, while positive statistics should not deteriorate too quickly. Choosing the right balance between these contradictory requirements is a challenge and we explore this topic further  in the evaluation section.  {\it Priya: we should address the one that works best such as exponentially weighted moving average in eval section - it doesn't belong here.}

>>>>

In order to distinguish between legitimate and malicious Interests, we leverage another unique feature of  NDN architecture --- guaranteed symmetric flow of Interest and Data packets. Since a Data packet takes the reverse path of the corresponding Interest packet, a router is guarateed to see if an Interest it forwarded resulted in a matching Data packet. The only exception being if Interest/Data packets were lost along the way due to congestion in the network.
Therefore, intermediate routers can classify all Interests that result in matching Data as legitimate, while the ones that timed out can be marked as malicious.\footnote{Recall that in order to maximize effect of the Interest flooding attack, an adversary expresses a large volume of junk Interests (see Section~\ref{sec:interest flooding}).}  
%Implications of other types of attacks are discussed in Section~\ref{sec:discussion}.}

This timeout-based differentiation method is reactive by nature: one cannot determine in advance if an Interest will result in a timeout or in Data being retrieved. However, routers can proactively maintain up-to-date statistics of Interests satisfaction ratios (number of forwarded versus number of satisfied Interests), and use this statistic to determine whether an incoming Interest must be forwarded or dropped. For example, maintaining independent Interest satisfaction ratio statistics for each incoming interface is sufficient to estimate whether an Interest received from a neighboring interface will result in a Data packet or a timeout. Statistics can also be kept at finer granularities such as per outgoing interface, per name prefix, etc. that can further improve the estimates. A router's goal is to penalize malicious Interests as soon as possible -- negative statistics should be allowed to build up fast, while positive statistics should not be allowed to deteriorate too quickly. Choosing the right balance between these contradictory requirements is a challenge and we explore this topic further  in the evaluation section.  {\it Priya: we should address the one that works best such as exponentially weighted moving average in eval section - it doesn't belong here.}

### 2012-12-19 algos/dynamic-limits.tex
Most of the evaluation section is semi-done.  I suspect I put too much details and discussion in the evalutions... Not sure
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@871 df1a90cd-247b-4364-bdaf-287a704eb499
More edits to chapters 4a, 4b
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@891 df1a90cd-247b-4364-bdaf-287a704eb499

Essentially, the probabilistic Interest accept algorithm divides the available forwarding tokens between interfaces proportionally to their satisfaction ratios.
Exactly the same effect can be achieved without using probabilistic methods, simply by enabling and enforcing additional Interest limits for each incoming interface, where the values of the limits directly depend on the interface satisfaction ratios.
Additionally, routers may need to explicitly announce these limits to their downstream neighbors, ensuring that any forwarded Interest from the downstream is actually getting through, resulting in genuine Interest satisfaction statistics.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\columnwidth]{dynamic-limits}
  \caption{Dynamic limits example: routers explicitly tell neighbors how many Interest can be for sure delivered to the Data producer}
  \label{fig:dynamic limits example}
\end{figure}

>>>>

Essentially, the probabilistic Interest accept algorithm divides the available forwarding tokens between interfaces in proportion to their Interest satisfaction ratios.
Given the drawbacks of this algorithm, the same effect can be achieved without using probabilistic methods. An alternate strategy is to enable and enforce additional Interest limits for each incoming interface, where the values of the limits directly depend on the interface satisfaction ratios.
Additionally, routers may need to explicitly announce these limits to their downstream neighbors, ensuring that any forwarded Interest from the downstream is actually getting through, resulting in genuine Interest satisfaction statistics.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{dynamic-limits}
  \caption{Dynamic limits example: routers explicitly tell neighbors how many Interest can be for sure delivered to the Data producer}
  \label{fig:dynamic limits example}
\end{figure}

### 2012-12-19 algos/probabilistic.tex
More edits to chapters 4a, 4b
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@891 df1a90cd-247b-4364-bdaf-287a704eb499

Apart from the Interest satisfaction statistics generation, there is a question how this statistics can be used to actually enforce prioritization and penalizing of Interests.
A straightforward way to achieve this enforcement is to consider Interest satisfaction rate as a direct probability to accept (forward) or reject an incoming Interest (see Pseudocode~\ref{alg:probabilistic model}).

@@

Parameter $\theta$ on line 5 of the Pseudocode~\ref{alg:probabilistic model} ensures that the probabilistic model is not applied when there a very small volume of Interests coming from the particular interface.
That is, while we want to deny service to the attackers, we also want give an opportunity for the users who do not abuse the network to regain their share of resources after temporary Data delivery failures.

A main drawback of the probabilistic Interests accept method is the fact that each router on the path make an independent decision on what to do with the Interest.
As a result of such independent decisions, the probability of legitimate Interests not to be dropped decreases quickly as the number of hops between the requester and Data grows, worsening the satisfaction statistics and resulting in further drops.
In example on Fig.~\ref{fig:flooding example}, the router A observes 50\% satisfaction rate for \texttt{eth1} and 0\% rate for \texttt{eth0}. 
At the same time, the router B sees the satisfaction rate for its \texttt{eth0} interface at 30\% level.
Next time a legitimate Interests arrives on the router A, it has 50\% chance to be forwarded further, and if forwarded, it has only $50\% \times 30\% = 15\%$ chance to be actually send towards the Data producer.
To prevent such overreacting, we need to enable some form of a gossiping protocol between neighboring NDN routers, in order to explicitly specify amount of Interests that each router is willing to forward.

>>>>

In addition to maintaining statistics on Interest satisfaction ratios, the next challenge is in using these statistics to penalize malicious Interests. A straightforward method to achieve this enforcement is to consider Interest satisfaction rate as a direct probability for accepting (forwarding) or rejecting an incoming Interest (see Pseudocode~\ref{alg:probabilistic model}).
%Apart from the Interest satisfaction statistics generation, there is a question how this statistics can be used to actually enforce prioritization and penalizing of Interests.

@@

Parameter $\theta$ on line 5 of the Pseudocode~\ref{alg:probabilistic model} ensures that the probabilistic model is not enforced when the volume of Interests arriving at a particular interface is small.
This step is critical -- while we want to drop Interests from attackers, we also want to provide an opportunity for legitimate users to regain their share of resources after temporary Data delivery failures due to  congestion.

A drawback of the probabilistic Interest acceptance method is that each router on the path makes an independent decision on whether to forward or drop the Interest. 
As a result of these independent decisions,  the probability of legitimate Interests being forwarded decreases rapidly as the number of hops between the content requester and producer grows; worsening the Interest satisfaction statistics and resulting in further drops.
In example on Fig.~\ref{fig:flooding example}, the router A observes 50\% satisfaction rate for \texttt{eth1} and 0\% rate for \texttt{eth0}. 
At the same time, router B observes a 30\% satisfaction rate for its \texttt{eth0} interface.
Next time a legitimate Interest arrives at router A, it has a 50\% chance of being forwarded further, and if forwarded, it has only a $50\% \times 30\% = 15\%$ probability of being forwarded further towards the Data producer. With each increasing hop in the network, the probability of being forwarded to the next hop decreases significantly. 
One way to prevent this overreaction and unfair penalization is to ensure that the decision taken at each router on whether to forward or drop the Interest is not independent of the decision taken at preceding routers. An explicit notification such as a gossip protocol between neighboring NDN routers that specifies the volume of Interests each router is willing to forward will likely address this issue.
